Novel genetic control of migratory diapause in Australian Monarch Butterflies

### 2023-06-16 paper/paper_draft.Rmd
added the NCBI files and the genotypes (zipped) updates to paper for molecular ecology

---
title: "Novel genetic control of migratory diapause in Australian Monarch Butterflies"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    reference_docx: "paper-files/style_ref.docx"
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
bibliography: "paper-files/citations.bib"
bst: "paper-files/pnas-new.bst"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
  - \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
  - \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
  - \usepackage[round]{natbib}
indent: true
sansfont: Times New Roman
fontsize: 12pt
always_allow_html: true
---

@@

Migratory species are acutely at risk from global anthropogenic environmental change, since they rely on multiple distinct ecosystems and consistent seasonal timings [@Both2010]. As a result, the loss of breeding, wintering, or any transitional habitats may all cause declines in migrants [@Robbins1989], as may shifts in resource phenology due to seasonal changes which can cause resource availability and species life-histories to fall out of sync [@Jones2010a]. It is therefore not surprising that widespread declines have been observed in many migratory taxa [@Limburg2009;@Robbins1989].

Population-level adaptability in migratory life-history should therefore be expected to increase long-term population viability by conferring resistance to phenological and environmental changes [@Pulido7341]. Migratory life-history adaptation can essentially occur along three distinct axis: 1) *temporal*, or adaptation that changes the phenology of migration; 2) *spatial*, or adaptation that alters migratory pathways and orientation; and 3) *residency*, or adaptation that shifts populations between migratory and resident life histories. Inter-population variation and adaptation in each of these pathways is well known from nature: temporal changes or variation in migration are well documented in many taxa, particularly in birds [@Rubolini2007; @Gordo2007];  as is short-stopping, extension of migratory pathways, or complete re-direction of migration [@Ambrosini2011; @LaSorte2007; @Visser2009]; and shifts towards residency or partial migration have been observed in birds [@Pulido7341; @Plummer2015; @Møller2014], butterflies [@Howard2010], mammals [@VanDerRee2006; @Fleming2003; @Jones2014a], and other taxa.

Adaptation via a transition to residency is of both particular interest and concern for the conservation of migratory species. While switching to a resident life history can allow species to avoid the increasing risks of migration and the intrinsically very large energetic costs of migration in general [@McWilliams2004; @Rankin1992; @Wikelski2003], there is reasonable concern that the "extinction" of migration may not be easily reversible in some species. This is concerning since migratory life-history variation is expected to increase population adaptability [@Pulido2007], and thus the permanent loss of migration could potentially compromise long-term population persistence. Beyond this, migratory species often provide critical ecosystem services [@lopez-hoffmanEcosystemServicesTransborder2017;@semmensAccountingEcosystemServices2011] and have cultural and/or ecological  significance [@burgerComplexityBioindicatorSelection2015;@schuetzCharacterizingCulturalNiches2019;@semmensQuantifyingEcosystemService2018;@reynoldsMigratorySpeciesEcological2011] that may not survive the transition to permanent residency.

The loss of migratory behavior is a major concern for monarch butterflies (*Danaus plexippus*) [@browerAnimalMigrationsEndangered1991;@satterfieldMigratoryMonarchsThat2018]. Specifically, North American migratory monarch populations have experienced severe declines over the last several decades, sparking a fear that those populations may go extinct in the near future [@Semmens2016]. However, many individual butterflies ostensibly from the North American migratory population are now year-round residents in some areas in the southern United States [@Howard2010], where they can survive due to the presence year-round of introduced tropical milkweed (*Asclepias curassavica*) [@bataldenPotentialChangesEastern2015], and there are wide-spread, non-migratory monarch populations that are not of any immediate conservation concern [@walkerDanausPlexippusErrata]. It is specifically, therefore, migration that is at risk in monarch butterflies, not the species as a whole. However, in addition to the damage to the long-term population persistence that could result from the loss of migration in North American populations (as discussed above), the annual monarch migration is publicly beloved and its loss would be culturally costly [@Gustafsson2015].

Interestingly, however, there a population of monarchs in Australia in which migratory behavior has apparently re-emerged after it was previously lost. This population, which was established sometime before 1871 [@Zalucki2004], is the product of a series of sequential introductions that crossed the Pacific Ocean from North America beginning sometime in the early 1840s after the introduction of their host plants enabled individuals to survive after being blown from the mainland [@Zalucki2004; @Pierce2014; @Hemstrom2022a]. Monarchs are not known to be migratory on any Pacific islands, and no migratory behavior was observed in Australia until the 1930s, when non-reproductive overwintering aggregations were observed in northern New South Wales [NSW, see @James2019]. These and other individuals were later shown to migrate hundreds of kilometers each year to their wintering grounds from more southerly, inland locations in NSW [@James2019;@jamesInductionReproductiveDormancy1983]. After over 90 years (and hundreds of generations) without any apparent migration, the behavior re-emerged in these populations. Understanding how this happened could provide crucial context around the "extinction" of migration in monarchs and other species.

The variance in diapause induction in Queensland monarchs allows for a direct search for genetic associations with this trait and thus for an examination of the genetic underpinnings of the re-evolution of migration in Australian monarchs. *Here, we used reduced representation sequencing to search for genes associated with migratory diapause in Queensland monarch butterflies. We found one strongly associated locus co-located with an E3 ubiquitin ligase gene. This gene has been previously identified as associated with intermolt transitions in silk moths (Bombyx mori), which is controlled by the same major hormone (juvenile hormone, hereafter JH) as reproductive diapause.* This gene had not previously been identified as associated with migration in North American monarch butterflies, and we did not find that any previously identified migratory genes were associated with diapause in our study.

## Background
Although the ancestral range of monarch butterflies is believed to be in North and Central America, they have fairly recently expanded into South America and the Caribbean, and very recently to several locations throughout the Atlantic and Pacific [@oplerFieldGuideWestern1999; @Zhan2014]. In North America, the species is well known for its unique, multi-generational migratory life history wherein individuals dispersing northward in the spring do so over three to four generations and then return to their wintering grounds in a single step [@malcolmSpringRecolonizationEastern1993; @cockrellTimeTemperatureLatitudinal1993].

In Southern Florida, the Caribbean, Central America, South America, and throughout most of their introduced range monarchs are non-migratory [@dockxDirectionalStabilizingSelection2007;@altizerPopulationsMonarchButterflies2010;@northMonarchButterflyBiology2004]. Residency in these populations appears to be a derived trait which has arisen multiple times after the species' split from *Danaus erripus*, their closest extant relative [@Zhan2014]. The Pacific expansion of monarchs is relatively recent, with historical records indicating that they established first in Hawaii in the 1840s, likely as a result of individuals blown in during storm events that were able to survive on recently introduced milkweeds [@Zalucki2004; @Pierce2014]. Genetic evidence is consistent with an introduction to Hawaii during this time period [@Hemstrom2022a]. As described above, they then spread across the Pacific and reached Australia in approximately 1871 [@Zalucki2004], and were recorded migrating again in Australia sometime in the 1930s [see Figure 1, @James2019].

@@

## Sequencing results
We incubated female monarch butterflies from a total of `r length(unique(na.omit(old.data$fam)))` clutches of eggs produced by females collected from a continously-breeding population from southern Queensland, Australia under a decreasing photoperiod following the methods of Freedman et al. [-@Freedman2018]. The resulting `r nrow(old.data)` females were phenotyped for oocyte development as a proxy for diapause status. `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "R", na.rm = TRUE)` butterflies were phenotyped as reproductively mature (e.g. not in diapause, and thus a non-migratory phenotype) and `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "B", na.rm = TRUE)`as reproductively immature (e.g. in diapause, and thus a migratory phenotype). We sequenced these individuals along with a additional `r sum(sample.meta$year == 2016, na.rm = TRUE)` butterflies raised under very similar conditions by Freedman et al. [-@Freedman2018], of which `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "R", na.rm = TRUE)` and `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "M", na.rm = TRUE)` were scored in that study as reproductively mature and immature, respectively.

We sequenced a total of 417,846,352 reads across all individuals, 98.8% of which mapped to the monarch reference genome. 244,443,414 reads were maintained after filtering (1,158,500 per sample, on average), encompassing `r prettyNum(L,big.mark=",",scientific=FALSE)` total sequenced bases. From this, we called `r prettyNum(nsnps(dat),big.mark=",",scientific=FALSE)` individual SNPs following filtering.

## Re-emergence of Diapaue Control in Australian Monarch Butterflies

We found that a region on Chromosome 11 containing the gene DPOGS208560 was strongly associated with diapause initiation in Monarch Butterflies in Queensland, Australia. This gene is an E3 ubiquitin ligase, a class of proteins broadly involved with proteosomal degradation of diverse substrates, including the circadian clock protein CRY2 [@xingSCFFBXL3UbiquitinLigase2013; @yooCompetingE3Ubiquitin2013], which is known to play a role in monarch photoperiodism including migratory responses to declining photoperiod [@iiamsPhotoperiodicClockRegulation2019]. Interestingly, a predicted homolog of this gene in silk moths has been previously found to be differentially expressed during larval molt-intermolt transitions [@Hu2016], which are generally controlled by fluctuations in JH and ecdysone [@Jindra2013]. JH is also known to be a major driver of reproductive development and diapause in monarchs [@Herman2001;@GreenII2019]. Taken together, it is not surprising that DPOGS208560 may participate in diapause control in monarchs.

What is surprising, however, is that this gene has not been previously identified as a candidate migration gene in monarchs. Specifically, DPOGS208560 was not among the 536 genes located in regions of the genome found to be significantly associated with contrasts between migration and residency in North and South American, Pacific, and European populations by Zhan et al. [-@Zhan2014], nor does it overlap with the genes suggested to be involved in photoperiodic responses by Iiams et al. [-@iiamsPhotoperiodicClockRegulation2019] or the genes associated with diapause termination in western North American monarchs [@GreenII2019]. There are at least two distinct explanations for this discrepancy: 1) the DPOGS208560 variant we observed is either present in North America but was not previously recognized as associated with migration or 2) is the result of a new mutation which occurred in Australian monarchs. The former case would imply that this allele was either maintained as monarchs passed through repeated population bottlenecks during their Pacific expansion or was secondarily re-introduced by immigrant individuals from North America (which would have likely been anthropogenic in origin).

Since we observed that *only*  the DPOGS208560 gene was associated with diapause in Australian monarchs, it seems unlikely that a secondary re-introduction of migratory monarch butterflies from North America could explain the re-emergence of migratory behavior in Australia, since there would be no reason for other known migratory associated genes to have not also made the jump back across the Pacific. Unfortunately, since the SNP loci that we located are not perfectly associated with diapause since only `r (gtable[1,1]/gtable[3,1])* 100` % of individuals phenotyped as without yolked oocytes and thus not in diapause carried the associated allele (compared to `r (gtable[1,2]/gtable[3,2])*100` % of reproductively mature individuals) and are likely to only be in linkage with any causal variation rather than the causal variants themselves, our work cannot currently delineate between the other two options (maintenance of variance at DPOGS208560 during the Pacific crossing or a new mutation). If the direct, causal loci were to be identified, it would then be possible to determine which gene versions are present in North America or across the Pacific and thus determine the origin of the DPOGS208560 diapause variant in Australia.

Note that it is unlikely that DPOGS208560 alone controls diapause induction under declining day-length in Australian monarchs. While our study features many *individuals*, it features only 32 maternal families and thus far fewer independant samples. Our power is therefore fairly limited, and it is entirely likely that we failed to detect many causal loci for diapause onset. This is consistent with the fact that our top associated SNP that overlaps DPOGS208560 was not perfectly associated with diapause, although that may also be due to imperfect linkage with the actual (and possibly un-sequenced) causal variants. Additional studies with larger sample sizes are therefore still needed in order to better understand the mechanisms underlying diapause induction in Australian monarch butterflies.

## Persistance of Migration in Monarchs and Beyond

Our work suggests that even if migration is lost in monarchs, it may not be lost forever. Despite the loss of allelic variance during the successive bottlenecks, monarch butterflies either maintained the genetic variation needed to migrate or secondarily re-aquired it via mutation. This gives us some hope that the contemporary loss of migration that we have observed in many migrants may be reversible over relatively short evolutionary timescales. At the very least, we can be assured that for monarchs in particular, migrants in Australia constitute a reservoir of migratory alleles that could potentially be tapped for North America if needed.

Alternatively, it is safe to assume that if migratory variance in Australian monarchs was maintained from North America, it is clear that this would only have been possible because selection was not actively acting against that variance. This is actually quite likely: monarchs in the Pacific are not exposed to substantial annual changes in day length or average temperature, and thus do not receive their main migratory cues. Environmentally triggered migratory variation, therefore, is probably selectively nearly neutral [although more fixed migratory variance, such as wing morphology, was likely selected against, driving the observed, repeated decreases in wing size and length observed in newly non-migratory monarch populations, see @Freedman2020], and is therefore more likely to be maintained in non-migratory populations. In contrast, a species with slower generation times and smaller effective population sizes would be unlikely re-evolve lost phenotypic variance on anything approaching a similar time frame, nor would they as easily hold on to phenotypic variance in the long run under neutral conditions. Nonetheless, our findings are still optimistic for the future of migration in monarch butterflies or wherever selection on migratory alleles is merely relaxed and not becoming increasingly negative.

# Conclusion

This study suggests that reproductive diapause in Australian monarch butterflies is influenced by a novel genetic mechanism via the DPOGS208560 gene, and that the previously identified migratory genetic elements were not associated with diapause in those populations. More research is needed to determine if the migratory variant of this gene is the result of ancestral variation maintained from North America or a new mutation. Regardless, our work is consistent with the hypothesis that proximate genetic mechanisms for migration are more likely to quickly re-emerge during a transition to a migratory life history than facultative mechanisms.

# Methods
## Sample Collection, Incubation, and Phenotyping

@@

In order to determine if diapause is part of a generalized migratory syndrome in Australia, we also briefly re-analyzed the connection between reproductive status and wing morphology in our 2016 samples [@Freedman2018]. In order to determine if wing shape or size was correlated with reproductive status, we fit a pair of linear mixed effect models with either with wing shape or size as response variables, the number of yolked oocytes as a fixed effect, and maternal family as a random effect using the R package nlme [@Pinheiro2007].

@@

@@

```{r Figure_3_print,echo=FALSE,warning=FALSE,fig.cap="Figure 3: Percentage of individuals in reproductive diapuase in each maternal family (A) and  GMMAT p-values for association between each SNP and diapause status across all individuals (B). SNPs with a p-value lower that 5 x 10^-5 are highlighted in red.",fig.width=8, fig.height=6, fig.align='center', fig.cap.style="Image Caption"}
grid.arrange(Figure3a + ggtitle("A"), Figure3b + ggtitle("B"), Figure3al, layout_matrix = matrix(c(1, 2, 3, NA), nrow = 2), widths = c(1, .1))

>>>>

---
title: "Novel genetic control of migratory diapause in Australian Monarch Butterflies"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    reference_docx: "paper-files/style_ref.docx"
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
bibliography: "paper-files/citations.bib"
csl: "paper-files/apa.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
  - \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
  - \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
  - \usepackage[round]{natbib}
indent: true
sansfont: Times New Roman
fontsize: 12pt
always_allow_html: true
---

@@

@@

# Abstract

Monarch Butterflies (Danaus plexippus) are a charismatic and culturally important North American butterfly species that are famous for their dramatic and unique migratory life history. While non-migratory populations of the species are widespread and not of particular conservation concern, migratory populations in North America have recently seen dramatic declines, prompting concern that the migratory life history of the species may vanish over the coming years as selective factors begin to favor residency over migration. However, one relatively new non-native monarch population in Australia rapidly re-acquired a migratory life history following hundreds of generations of residency and successive bottlenecks as the species stepping-stoned across the Pacific during the late 1800's and early 1900's. The genetic process by which this happened is not currently known. We raised and sequenced individuals from Queensland, Australia and found a novel association between reproductive diapause, a key migratory trait, and a genetic locus that has not previously been identified to be involved in migration in North America. This may imply that the species might be able to rapidly re-evolve migration in other places as well and thus provides some hope that the feared loss of migration in North American monarchs may not be permanent.

Since they rely on multiple distinct ecosystems and consistent seasonal timings, migratory species are acutely at risk from global anthropogenic environmental change [@Both2010; @chowdhurySeasonalSpatialDynamics2021].The loss of breeding, wintering, or any transitional habitats may all cause declines in migrants [@Robbins1989], as may shifts in resource phenology due to seasonal changes which can cause resource availability and species life-histories to fall out of synchrony [@Jones2010a]. It is therefore not surprising that widespread declines have been observed in many migratory taxa [@Limburg2009;@Robbins1989;@rungeConservingMobileSpecies2014].

Population-level adaptability in migratory life-history should therefore be expected to increase long-term population viability by conferring resistance to phenological and environmental changes [@Pulido7341]. Migratory life-history adaptation can essentially occur along three distinct axes: 1) *temporal*, or adaptation that changes the phenology of migration; 2) *spatial*, or adaptation that alters migratory pathways and orientation; and 3) *residency*, or adaptation that shifts populations between migratory and resident life histories. Inter-population variation and adaptation in each of these pathways is well known from nature: temporal changes or variation in migration are well documented in many taxa including as birds [@Rubolini2007; @Gordo2007; @chuCommunityScienceReveals2022; @prytulaRapidAdvancementSpring2022; @pinszkeLongtermChangesAutumn2023; @mondain-monvalFlywayscaleAnalysisReveals2021], fish [@kovachGeneticChangeEarlier2012], mammals [@shuertDecadalMigrationPhenology2022], and butterflies [@culbertsonLongtermMonitoringIndicates2022; @fitchettAdvanceTimingAnnual2022];  as is short-stopping, extension of migratory pathways, or complete re-direction of migration [@Ambrosini2011; @LaSorte2007; @Visser2009; @sonnleitnerRapidShiftsMigration2022;@rushingMigratoryBehaviorWinter2020;@dufourYellowbrowedWarblerPhylloscopus2022;@dufourNewWestwardMigration2021;@jiaShiftingMigrationRoute2021]; and shifts towards residency or partial migration have been observed and/or predicted in birds [@Pulido7341; @Plummer2015; @Møller2014;@haranPatternsPartialMigration2021; @falchiInterIntrapopulationVariability2023;@bonnetlebrunEffectsUrbanizationBird2020], butterflies [@Howard2010], mammals [@VanDerRee2006; @Fleming2003; @Jones2014a], fish [@adamsOpinionPieceEvolutionary2022], and other taxa.

Adaptation via a transition to residency is of both particular interest and concern for the conservation of migratory species. While switching to a resident life history can allow species to avoid the increasing risks of migration and the intrinsically large energetic costs of migration in general [@McWilliams2004; @Rankin1992; @Wikelski2003], there is reasonable concern that the "extinction" of migration may not be easily reversible in some species. This is concerning since migratory life-history variation is expected to increase population adaptability [@Pulido2007], and thus the permanent loss of migration could potentially compromise long-term population persistence. Beyond this, migratory species often provide critical ecosystem services [@lopez-hoffmanEcosystemServicesTransborder2017;@semmensAccountingEcosystemServices2011] and have cultural and/or ecological  significance [@burgerComplexityBioindicatorSelection2015;@schuetzCharacterizingCulturalNiches2019;@semmensQuantifyingEcosystemService2018;@reynoldsMigratorySpeciesEcological2011] that may not survive the transition to permanent residency.

The loss of migratory behavior is a major concern for monarch butterflies (*Danaus plexippus*) [@browerAnimalMigrationsEndangered1991;@satterfieldMigratoryMonarchsThat2018]. Specifically, North American migratory monarch populations have experienced severe declines over the last several decades, sparking a fear that those populations may go extinct in the near future [@Semmens2016]. However, many individual butterflies ostensibly from the North American migratory population are now year-round residents in some areas in the southern United States [@Howard2010], where they can survive due to the presence year-round of introduced tropical milkweed (*Asclepias curassavica*) [@bataldenPotentialChangesEastern2015], and there are wide-spread, non-migratory monarch populations that are not of any immediate conservation concern [@walkerDanausPlexippusErrata]. While the species as a whole may therefore not be at risk, their migratory behavior almost certainly is. In addition to the damage to the long-term population persistence that could result from the loss of migration in North American populations (as discussed above), the loss of the publicly beloved annual monarch migration and concomitant overwintering would be culturally costly [@Gustafsson2015].

Interestingly, however, there is a population of monarchs in Australia in which migratory behavior has apparently re-emerged after it was previously lost. This population, which was established sometime before 1871 [@Zalucki2004], is the product of a series of sequential introductions that crossed the Pacific Ocean from North America beginning sometime in the early 1840s after the introduction of their host plants enabled individuals to survive after being blown from the mainland [Figure 1, see @Zalucki2004; @Pierce2014; @Hemstrom2022a]. Monarchs are not known to be migratory on any Pacific islands, and no migratory behavior was observed in Australia until the 1930s, when non-reproductive overwintering aggregations were observed in northern New South Wales [NSW, see @James2019]. These and other individuals were later shown to migrate hundreds of kilometers each year to their wintering grounds from more southerly, inland locations in NSW [@James2019;@jamesInductionReproductiveDormancy1983]. After over 90 years (and hundreds of generations) without any apparent migration, the behavior re-emerged in these populations. Understanding how this happened could provide crucial context around the "extinction" of migration in monarchs and other species.

The variance in diapause induction in Queensland monarchs allows for a direct search for genetic associations with this trait and thus for an examination of the genetic underpinnings of the re-evolution of migration in Australian monarchs. Here, we used reduced representation sequencing to search for genes associated with migratory diapause in Queensland monarch butterflies. We found one strongly associated locus co-located with an E3 ubiquitin ligase gene. This gene has been previously identified as associated with intermolt transitions in silk moths (Bombyx mori), which is controlled by the same major hormone (juvenile hormone, hereafter JH) as reproductive diapause. This gene had not previously been identified as associated with migration in North American monarch butterflies, and we did not find that any previously identified migratory genes were associated with diapause in our study.

# Materials and Methods
## Sample Collection, Incubation, and Phenotyping

@@

To determine if diapause is part of a generalized migratory syndrome in Australia we also briefly re-analyzed the connection between reproductive status and wing morphology in our 2016 samples [@Freedman2018]. To determine if wing shape or size was correlated with reproductive status, we fit a pair of linear mixed effect models with either with wing shape or size as response variables, the number of yolked oocytes as a fixed effect, and maternal family as a random effect using the R package nlme [@Pinheiro2007].

## Background
Although the ancestral range of monarch butterflies is believed to be in North and Central America, they have fairly recently expanded into South America and the Caribbean, and very recently to several locations throughout the Atlantic and Pacific [@oplerFieldGuideWestern1999; @Zhan2014; @haegerMonarchsAtlanticOcean2015]. In North America, the species is well known for its unique, multi-generational migratory life history wherein individuals dispersing northward in the spring do so over three to four generations and then return to their wintering grounds in a single step [@malcolmSpringRecolonizationEastern1993; @cockrellTimeTemperatureLatitudinal1993].

In Southern Florida, the Caribbean, Central America, South America, and throughout most of their introduced range monarchs are non-migratory [@dockxDirectionalStabilizingSelection2007;@altizerPopulationsMonarchButterflies2010;@northMonarchButterflyBiology2004]. Residency in these populations appears to be a derived trait which has arisen multiple times after the species' split from *Danaus eripus*, their closest extant relative [@Zhan2014]. The Pacific expansion of monarchs is relatively recent, with historical records indicating that they established first in Hawaii in the 1840s, likely as a result of individuals blown in during storm events that were able to survive on recently introduced milkweeds [@Zalucki2004; @Pierce2014]. Genetic evidence is consistent with an introduction to Hawaii during this time period [@Hemstrom2022a]. As described above, they then spread across the Pacific and reached Australia in approximately 1871 [@Zalucki2004], and were recorded migrating again in Australia sometime in the 1930s [see Figure 1, @James2019].

## Sequencing results
We incubated female monarch butterflies from a total of `r length(unique(na.omit(old.data$fam)))` clutches of eggs produced by females collected from a continuously-breeding population from southern Queensland, Australia under a decreasing photoperiod following the methods of Freedman et al. [-@Freedman2018]. The resulting `r nrow(old.data)` females were phenotyped for oocyte development as a proxy for diapause status. Butterflies were phenotyped as reproductively mature (n = `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "R", na.rm = TRUE)` not in diapause and thus with a non-migratory phenotype) or as reproductively immature (n = `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "M", na.rm = TRUE)` in diapause and thus with a migratory phenotype). We sequenced these individuals along with an additional `r sum(sample.meta$year == 2016, na.rm = TRUE)` butterflies raised under very similar conditions by Freedman et al. [-@Freedman2018], of which `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "R", na.rm = TRUE)` and `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "M", na.rm = TRUE)` were scored in that study as reproductively mature and immature, respectively.

We sequenced a total of 417,846,352 reads across all individuals, 98.8% of which mapped to the monarch reference genome and 244,443,414 of which were maintained after filtering (1,158,500 per sample, on average), encompassing `r prettyNum(L,big.mark=",",scientific=FALSE)` total sequenced bases. From this, we called `r prettyNum(nsnps(dat),big.mark=",",scientific=FALSE)` individual SNPs following filtering.

## Re-emergence of Diapause Control in Australian Monarch Butterflies

We found that a region on Chromosome 11 containing the gene DPOGS208560 was strongly associated with diapause initiation in Monarch Butterflies in Queensland, Australia. This gene is an E3 ubiquitin ligase, a class of proteins broadly involved with proteosomal degradation of diverse substrates, including the circadian clock protein CRY2 [@xingSCFFBXL3UbiquitinLigase2013; @yooCompetingE3Ubiquitin2013], which is known to play a role in monarch photoperiodism including migratory responses to declining photoperiod [@iiamsPhotoperiodicClockRegulation2019]. Interestingly, a predicted homolog of this gene in silk moths has been previously found to be differentially expressed during larval molt-intermolt transitions [@Hu2016], which are generally controlled by fluctuations in JH and ecdysone [@Jindra2013]. JH is known to be a major driver of reproductive development and diapause in monarchs [@Herman2001;@GreenII2019]. Taken together, it is not surprising that DPOGS208560 may participate in diapause control in monarchs.

What is surprising, however, is that this gene has not been previously identified as a candidate migration gene in monarchs. Specifically, DPOGS208560 was not among the 536 genes located in regions of the genome found to be significantly associated with contrasts between migration and residency in North and South American, Pacific, and European populations by Zhan et al. [-@Zhan2014], nor does it overlap with the genes suggested to be involved in photoperiodic responses by Iiams et al. [-@iiamsPhotoperiodicClockRegulation2019] or the genes associated with diapause termination in western North American monarchs [@GreenII2019]. There are at least two distinct explanations for this difference: 1) the DPOGS208560 variant we observed is either present in North America but was not previously recognized as associated with migration or 2) is the result of a new mutation which occurred in Australian monarchs. The former case would imply that this allele was either maintained as monarchs passed through repeated population bottlenecks during their Pacific expansion or was secondarily re-introduced by immigrant individuals from North America (which would have likely been anthropogenic in origin).

Since we observed that *only*  the DPOGS208560 gene was associated with diapause in Australian monarchs, it seems unlikely that a secondary re-introduction of migratory monarch butterflies from North America could explain the re-emergence of migratory behavior in Australia, since there would be no reason for other known migratory associated genes to have not also made the jump back across the Pacific. Unfortunately, the SNP loci that we located are not perfectly associated with diapause since only `r (gtable[1,1]/gtable[3,1])* 100` % of individuals phenotyped as not in diapause carried the associated allele (compared to `r (gtable[1,2]/gtable[3,2])*100` % of individuals phenotyped as reproductively mature) and are likely to only be in linkage with any causal variation rather than the causal variants themselves. Our work cannot therefore currently delineate between the other two options (maintenance of variance at DPOGS208560 during the Pacific crossing or a new mutation). If the direct, causal loci were to be identified, it would then be possible to determine which gene versions are present in North America or across the Pacific and thus determine the origin of the DPOGS208560 diapause variant in Australia.

Note that it is unlikely that DPOGS208560 alone controls diapause induction under declining day-length in Australian monarchs. While our study features many *individuals*, it features only 32 maternal families and thus far fewer independent samples. Our power is therefore limited, and it is entirely likely that we failed to detect many causal loci for diapause onset. This is consistent with the fact that our top associated SNP that overlaps DPOGS208560 was not perfectly associated with diapause, although that may also be due to imperfect linkage with the actual (and possibly un-sequenced) causal variants. Additional studies with larger sample sizes are therefore still needed to better understand the mechanisms underlying diapause induction in Australian monarch butterflies.

## Persistence of Migration in Monarchs and Beyond

Our work suggests that even if migration is lost in monarchs, it may be temporary. Despite the loss of allelic variance during the successive bottlenecks, monarch butterflies either maintained the genetic variation needed to migrate or secondarily re-acquired it via mutation. This gives us some hope that the contemporary loss of migration that we have observed in many migrants may be reversible over relatively short evolutionary timescales. At the very least, we can be assured that for monarchs in particular, migrants in Australia constitute a reservoir of migratory alleles that could potentially be tapped for North America if needed.

Alternatively, it is safe to assume that if migratory variance in Australian monarchs was maintained from North America this would only have been possible because selection was not actively acting against that variance. This is actually quite likely: monarchs in the Pacific are not exposed to substantial annual changes in day length or average temperature, and thus do not receive their main migratory cues. Environmentally triggered migratory variation, therefore, is probably selectively nearly neutral in the Pacific. In contrast, individually fixed migratory variance, such as wing morphology, was likely selected against across the Pacific, thus driving the observed, repeated decreases in wing size and length observed in newly non-migratory monarch populations [@Freedman2020]. Environmentally triggered migratory variation is therefore more likely type of migratory variation to be maintained in non-migratory populations.

In contrast to monarchs, however, a species with slower generation times and smaller effective population sizes would be unlikely re-evolve lost phenotypic variance on anything approaching a similar time frame, nor would they as easily hold on to phenotypic variance in the long run under neutral conditions. Nonetheless, our findings are still optimistic for the future of migration in monarch butterflies or in other species where selection on migratory alleles is merely relaxed and not becoming increasingly negative.

# Conclusion
This study suggests that reproductive diapause in Australian monarch butterflies is influenced by a novel genetic mechanism via the DPOGS208560 gene, and that the previously identified migratory genetic elements were not associated with diapause in this populations. Research to determine if the migratory variant of this gene is the result of ancestral variation maintained from North America or a new mutation is needed. Regardless, our work is consistent with the hypothesis that proximate genetic mechanisms for migration are more likely to quickly re-emerge during a transition to a migratory life history than facultative mechanisms.

# Acknowledgements and Funding Sources
We would like to thank Dr. Shawan Chowdhury for his assistance with Monarch feeding and maintenance during the rearing process and Dr. Sean O’Rourke for assistance during library preparation and sequencing. Travel and sequencing funding was provided by the Jastro-Shields Graduate Research Award and the UC Davis Department of Animal Science. MF was funded by NSF Postdoctoral Fellowship Award 2010658.

@@

# Data Availability and Benefits-Sharing
Data Availability: The raw sequencing data produced for this study are available from the NCBI under accession number (to be provided). The filtered genotypes and the scripts used to produce both them and this paper are available at https://github.com/hemstrow/aus_monarchs. The exact script used to produce this paper from the genotypic data, including all non-map figures, is available at https://github.com/hemstrow/aus_monarchs/blob/master/paper/paper_draft.Rmd.

Benefit-Sharing: This paper represents the continuation of a collaboration between researchers in the United States and Australia. All major contributors are listed as co-authors. This paper also extends our knowledge of critical phenotypic variance in a newly-listed endangered species. All of the data produced in this paper are shared with the public at the repositories above.

# Author Contributions
WH and MM designed the study with input from MF and MZ. WH collected, raised, and phenotyped individuals. MZ provided, lab space, equipment, and expertise for each of those steps. WH analyzed the data and wrote the manuscript. MF, MZ, and MM provided feedback on the analysis and manuscript.

@@

```{r Figure_3_print,echo=FALSE,warning=FALSE,fig.cap="Figure 3: Percentage of individuals in reproductive diapause in each maternal family (A) and  GMMAT p-values for association between each SNP and diapause status across all individuals (B). SNPs with a p-value lower that 5 x 10^-5 are highlighted in red.",fig.width=8, fig.height=6, fig.align='center', fig.cap.style="Image Caption"}
grid.arrange(Figure3a + ggtitle("A"), Figure3b + ggtitle("B"), Figure3al, layout_matrix = matrix(c(1, 2, 3, NA), nrow = 2), widths = c(1, .1))
